Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex
Initial draft of future work section.

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  Given more time, we are confident that
this could be completed however.  So far, we have proven the first few lemmas
described in Section~\ref{sec:proof}.  This was much more difficult than the
short natural language proofs would suggest, due to a combination of the
stricter format required by Coq and the number of items that are elided in the
proofs because they were considered trivial or obvious.

@@

\todo{Talk about stuff we still have to do for this to be a complete project.
Talk about expected problems and required techniques.  Talk about what we could
do once this project is completed and we could use it for other projects.}

>>>>

As mentioned above, we unfortunately were unable to complete our noninterference
proof in Coq by the time of this report.  Given more time, we are confident that
this could be completed however.  So far, we have proven the first few lemmas
described in Section~\ref{sec:proof} and defined relations for the operational
semantics and type system as presented in Section~\ref{sec:lang}.  This was much
more difficult than the short natural language proofs would suggest, due to a
combination of the stricter format required by Coq and the number of items that
are elided in the proofs because they were considered trivial or obvious.

@@

It is our hope that we could continue this work to complete a Coq proof of
noninterference for our language.  This would require proving the lemmas and
theorem in Section~\ref{sec:proof} along with various support lemmas and
definitions for things that are ``obvious'' in the natural language proofs.
It's unclear how long the remainder of the proof would take, but we are certain
that it is just a matter of time and effort.

Once a full Coq proof of noninterference exists, a next step would be to
integrate this proof and all of the related definitions into a verified type
checking program, which could be intergrated into a verified compiler.  We
already know that a full verified compiler has been done for C, and this gives
us hope that it would be possible to implement something similar for a language
like Jif.

With a verified compiler for such a language, it would be possible to implement
a compiler that produced proofs of noninterference to be carried around with the
compiled program.  This would be interesting in distributed settings where we
would like to ship code around to others to run on their own machines.  There is
already a paper on the subject of mobile code for the Fabric
language~\cite{liu2009fabric}, a distributed computing variant of Jif,
discussing the problem of ``mobile code'', which operates by having the client
compile the code themselves~\cite{arden2012sharing}.  A proof-carrying code
system could possibly remove the need for the target machine to compile the code
themselves in order to ensure trust requirements are met.

Another use of a verified compiler would be the ability to verify extensions to
the language.  For example, the Jif and Fabric programming languages are
actively worked on and has seen multiple extensions over the years.  Although
it's pretty certain extensions to these languages that are published are
correct, requiring these extensions to be added to a verified compiler would
help ensure that the language updates will not invalidate pre-existing security
properties of the language, because adding the extension would require adding
code to cover the additional cases in the proofs.

\todo{more.}
